---
title: "Incremental Mixture Importance Sampling"
author: "DARTH Workgroup"
date: "October 14, 2018"
output:
    html_document:
        toc: true
        depth: 3
        number_sections: true
        theme: united
        toc_float: true
        highlight: tango
---

Bayesian methods allow us to quantify the uncertainty in the calibrated parameters even in the presence of non-identifiability.

To conduct a Bayesian calibration of the three-satte model, we use the incremental mixture importance sampling (IMIS) algorithm. 

# Setup

## Packages

```{r message=FALSE}
# calibration functionality
library(lhs)
library(IMIS)
library(matrixStats) # package used for sumamry statistics
library(modeest)     # package used to estimate mode

# visualization
library(plotrix)
library(psych)
```

## Targets

```{r}
load("CRSTargets.RData")

# Plot the targets

# TARGET 1: Survival ("Surv")
plotrix::plotCI(x = CRS.targets$Surv$Time, y = CRS.targets$Surv$value, 
                ui = CRS.targets$Surv$ub,
                li = CRS.targets$Surv$lb,
                ylim = c(0, 1), 
                xlab = "Time", ylab = "Pr Survive")

```


## Model

```{r}
source("markov_crs.R") # creates the function markov_crs()
```

# Calibration

We're going to use the `IMIS()` function within the `IMIS` package. This function needs three functions to be defined in the environment: `prior()`, `likelihood()`, and `sample.prior()`. In the description of the `IMIS` package authors: "prior(x) calculates prior density of x, likelihood(x) calculates the likelihood of x, and sample.prior(n) draws n samples from the prior distribution". So, we'll define each of those.


## Sample Prior

The following defines a *uniform* prior on each of the parameters.

```{r}
# names and number of input parameters to be calibrated
param.names <- c("p.Mets","p.DieMets")
n.param <- length(param.names)

# range on input search space
lb <- c(p.Mets = 0.04, p.DieMets = 0.04) # lower bound
ub <- c(p.Mets = 0.16, p.DieMets = 0.16) # upper bound

sample.prior <- function(n.samp){
  m.lhs.unit   <- lhs::randomLHS(n = n.samp, k = n.param)
  m.param.samp <- matrix(nrow = n.samp, ncol = n.param)
  colnames(m.param.samp) <- param.names
  for (i in 1:n.param){
    m.param.samp[, i] <- qunif(m.lhs.unit[,i],
                               min = lb[i],
                               max = ub[i])
  }
  return(m.param.samp)
}

# view resulting parameter set samples
psych::pairs.panels(sample.prior(1000))
```

Note that we could define a different prior on the parameters instead of uniform, like
```
m.param.samp[, i] <- qbeta(m.lhs.unit[,i],
                           shape1 = 1,
                           shape2 = 1))
```


## Prior density 

The following function evaluates the prior at `v.params`.

```{r}
f_log_prior <- function(v.params){
  if(is.null(dim(v.params))) { # If vector, change to matrix
    v.params <- t(v.params) 
  }
  n.samp <- nrow(v.params)
  colnames(v.params) <- param.names
  lprior <- rep(0, n.samp)
  for (i in 1:n.param){
    lprior <- lprior + dunif(v.params[, i],
                             min = lb[i],
                             max = ub[i], 
                             log = T)
  }
  return(lprior)
}
```

The result of this function is a single number - the log of the prior density at `v.params`

```{r}
v.params.test <- c("p.Mets" = 0.1, "p.DieMets" = 0.2)
f_log_prior(v.params = v.params.test)
```

We exponentiate to get the prior density at `v.params`.

```{r}
prior <- function(v.params) { 
  exp(f_log_prior(v.params)) 
}
```

## Likelihood

The final piece for `IMIS` is the likelihood function. As before, we calculate the log-likelihood function and exponentiate to get `likelihood()`.

```{r}

# number of calibration targets
target.names <- c("Surv")
n.target     <- length(target.names)

f_llik <- function(v.params){
  # par_vector: a vector (or matrix) of model parameters 
  if(is.null(dim(v.params))) { # If vector, change to matrix
    v.params <- t(v.params) 
  }
  n.samp <- nrow(v.params)
  v.llik <- matrix(0, nrow = n.samp, ncol = n.target) 
  llik.overall <- numeric(n.samp)
  for(j in 1:n.samp) { # j=1
    jj <- tryCatch( { 
      ###   Run model for parameter set "v.params" ###
      model.res <- markov_crs(v.params[j, ])
      
      ###  Calculate log-likelihood of model outputs to targets  ###
      # Survival ("Surv")
      v.llik[j, 1] <- sum(dnorm(x = CRS.targets$Surv$value,
                                mean = model.res$Surv,
                                sd = CRS.targets$Surv$se,
                                log = T))
      
      # OVERALL 
      llik.overall[j] <- sum(v.llik[j, ])
    }, error = function(e) NA) 
    if (is.na(jj)) { llik.overall <- -Inf }
  } # End loop over sampled parameter sets
  # return LLIK
  return(llik.overall)
}
```

The results are a single number:
```{r}
f_llik(v.params = v.params.test)
```

Now, define the likelihood:
```{r}
likelihood <- function(v.params){ 
  exp(f_llik(v.params)) 
}
likelihood(v.params = v.params.test)
```


## Log-posterior
```{r}
f_log_post <- function(v.params) { 
  lpost <- f_log_prior(v.params) + f_llik(v.params)
  return(lpost) 
}
f_log_post(v.params = v.params.test)
```

## Calibrate!

```{r}
# number of resamples
n.resamp <- 1000

# run IMIS
fit.imis <- IMIS(B = 1000, # the incremental sample size at each iteration of IMIS.
                 B.re = n.resamp, # the desired posterior sample size
                 number_k = 10, # the maximum number of iterations in IMIS.
                 D = 0)
```

The posterior distribution is stored in `fit.imis$resample`:

```{r}
# obtain posterior
m.calib.post <- fit.imis$resample
head(m.calib.post)
```

Let's look at the results!

```{r}
# Plot the 1000 draws from the posterior
plot(m.calib.post,
     xlim = c(lb[1], ub[1]), ylim = c(lb[2], ub[2]),
     xlab = param.names[1], ylab = param.names[2])

# Plot the 1000 draws from the posterior with marginal histograms
psych::pairs.panels(m.calib.post)
```

Other posterior statistics - the mean, median, and mode. Note that the mode gives us the *maximum a posteriori* estimate:

```{r warning=FALSE}
# Compute posterior mean
v.calib.post.mean <- colMeans(m.calib.post)
v.calib.post.mean

# Compute posterior median and 95% credible interval
m.calib.post.95cr <- matrixStats::colQuantiles(m.calib.post, probs = c(0.025, 0.5, 0.975))
m.calib.post.95cr

# Compute posterior mode
v.calib.post.mode <- apply(m.calib.post, 2, function(x) as.numeric(modeest::mlv(x)[1]))
v.calib.post.mode

# compute maximum a posteriori
v.calib.like <- likelihood(m.calib.post)
map <- m.calib.post[which.max(v.calib.like), ]
```


## Calibrated model predictions vs. targets

### Mode
```{r}
v.out.post.mode <- markov_crs(v.calib.post.mode)

# TARGET 1: Survival ("Surv")
plotrix::plotCI(x = CRS.targets$Surv$Time, y = CRS.targets$Surv$value, 
                ui = CRS.targets$Surv$ub,
                li = CRS.targets$Surv$lb,
                ylim = c(0, 1), 
                xlab = "Time", ylab = "Pr Survive")
grid()
for (i in 1:nrow(m.calib.post)){
  mod_output <- markov_crs(m.calib.post[i, ])
  lines(x = CRS.targets$Surv$Time, 
        y = mod_output$Surv,
        col = "darkorange",
        lwd = 0.1)
}
lines(x = CRS.targets$Surv$Time, 
       y = v.out.post.mode$Surv,
      col = "dodgerblue",
      lwd = 2)
```


### MAP
```{r}
v.out.post.map <- markov_crs(map)

# TARGET 1: Survival ("Surv")
plotrix::plotCI(x = CRS.targets$Surv$Time, y = CRS.targets$Surv$value, 
                ui = CRS.targets$Surv$ub,
                li = CRS.targets$Surv$lb,
                ylim = c(0, 1), 
                xlab = "Time", ylab = "Pr Survive")
grid()
for (i in 1:nrow(m.calib.post)){
  mod_output <- markov_crs(m.calib.post[i, ])
  lines(x = CRS.targets$Surv$Time, 
        y = mod_output$Surv,
        col = "darkorange",
        lwd = 0.1)
}
lines(x = CRS.targets$Surv$Time, 
       y = v.out.post.map$Surv,
      col = "dodgerblue",
      lwd = 2)
```


# Return to Main Page

<a href="https://caleb-easterly.github.io/calibSMDM2018-materials/">Back to main page</a>
